---
title: "Predictions using the Weight Lifting Exercises Dataset"
Author: "Yassine Hachaichi"
output: html_notebook
---
1. Explore the data, especially focussing on the two paramaters we are interested in
2. Model selection, where we try different models to help us answer our questions
3. Model examination, to see wether our best model holds up to our standards
4. A Conclusion where we answer the questions based on the data
5. Predicting the classification of the model on test set

## Importing data and exploation
```{r}
training <- read.csv("./pml-training.csv")
testing <- read.csv("./pml-testing.csv")
```

```{r}
dim(training)
```

```{r}
head(training)
```

## Cleaning data
```{r}
maxNAPerc = 20
maxNACount <- nrow(training) / 100 * maxNAPerc
removeColumns <- which(colSums(is.na(training) | training=="") > maxNACount)
training.cleaned01 <- training[,-removeColumns]
testing.cleaned01 <- testing[,-removeColumns]
```

```{r}
removeColumns <- grep("timestamp", names(training.cleaned01))
training.cleaned02 <- training.cleaned01[,-c(1, removeColumns )]
testing.cleaned02 <- testing.cleaned01[,-c(1, removeColumns )]
```

```{r}
classeLevels <- levels(training.cleaned02$classe)
training.cleaned03 <- data.frame(data.matrix(training.cleaned02))
training.cleaned03$classe <- factor(training.cleaned03$classe, labels=classeLevels)
testing.cleaned03 <- data.frame(data.matrix(testing.cleaned02))
```

```{r}
training.cleaned <- training.cleaned03
testing.cleaned <- testing.cleaned03
```

```{r}
set.seed(19791108)
library(caret)
classeIndex <- which(names(training.cleaned) == "classe")
partition <- createDataPartition(y=training.cleaned$classe, p=0.75, list=FALSE)
training.subSetTrain <- training.cleaned[partition, ]
training.subSetTest <- training.cleaned[-partition, ]
```
## Feature correlations

```{r}
correlations <- cor(training.subSetTrain[, -classeIndex], as.numeric(training.subSetTrain$classe))
bestCorrelations <- subset(as.data.frame(as.table(correlations)), abs(Freq)>0.3)
bestCorrelations
```

## Some graphical representations

```{r}
library(Rmisc)
library(ggplot2)
p1 <- ggplot(training.subSetTrain, aes(classe,pitch_forearm)) +
geom_boxplot(aes(fill=classe))
p2 <- ggplot(training.subSetTrain, aes(classe, magnet_arm_x)) +
geom_boxplot(aes(fill=classe))
multiplot(p1,p2,cols=2)
```

## The correlations heatmap

```{r}
library(corrplot)
correlationMatrix <- cor(training.subSetTrain[, -classeIndex])
highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.9, exact=TRUE)
excludeColumns <- c(highlyCorrelated, classeIndex)
corrplot(correlationMatrix, method="color", type="lower", order="hclust", tl.cex=0.70, tl.col="black", tl.srt = 45, diag = FALSE)
```


## Classification methods
## Random Forest
```{r}
library(rpart)
library(rpart.plot)
library(rattle)
training <- read.csv("./pml-training.csv")
testing <- read.csv("./pml-testing.csv")
label <- createDataPartition(training$classe, p = 0.7, list = FALSE)
train <- training[label, ]
test <- training[-label, ]
```
```{r}
NZV <- nearZeroVar(train)
train <- train[ ,-NZV]
test <- test[ ,-NZV]
label <- apply(train, 2, function(x) mean(is.na(x))) > 0.95
train <- train[, -which(label, label == FALSE)]
test <- test[, -which(label, label == FALSE)]
train <- train[ , -(1:5)]
test <- test[ , -(1:5)]
```

```{r}

library(caret)
set.seed(13908)
control <- trainControl(method = "cv", number = 3, verboseIter=FALSE)
modelRF <- train(classe ~ ., data = train, method = "rf", trControl = control)
modelRF$finalModel
```


```{r}
predictRF <- predict(modelRF, test)
confMatRF <- confusionMatrix(predictRF, test$classe)
confMatRF
```
## Gradient Boosting
```{r}
library(rpart)
library(rpart.plot)
library(rattle)
```

```{r}
control <- trainControl(method = "repeatedcv", number = 5, repeats = 1, verboseIter = FALSE)
modelGBM <- train(classe ~ ., data = train, trControl = control, method = "gbm", verbose = FALSE)
modelGBM$finalModel
```

```{r}
predictGBM <- predict(modelGBM, test)
confMatGBM <- confusionMatrix(predictGBM, test$classe)
confMatGBM
```
